Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package

### 2020-10-20 040-monte-carlo-evaluation.Rmd
dir create moved outside of loop and set recursive
some directories did not exist, and so building the book would fail

```{r LoadLib040, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
## reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

res <- data.frame()

for (i in 1:nrow(sim_pars)) {
    type <- sim_pars$type[i]
    cores <- sim_pars$cores[i]
    n <- sim_pars$n[i]
    adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
    tmp_bounds <- cor_bounds(margins, type = type)
    cor_lo <- tmp_bounds$lower[1,2] + eps
    cor_hi <- tmp_bounds$upper[1,2] - eps
    cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
    ## rho = cor_seq[1]
    for (rho in cor_seq) {
        Rho <- matrix(rho, 2, 2)
        diag(Rho) <- 1.0
        ## Rho <- convertCor( rho = Rho, from = type, to = 'pearson' )
        time_data <- system.time({
            ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
            ## use rvec to test our software
            x <- rvec(n = n,
                      rho = Rho,
                      margins = margins,
                      type = type,
                      cores = cores)
        })
        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "1CORE",
            "-lib", "bigsimr"
        )
        ## save sims in their own subdirectory
        dir.create("./results/sims_norm", showWarnings = FALSE)
        saveRDS(x, file = paste0("./results/sims_norm/", id, ".rds"))

@@

res <- data.frame()
for (i in 1:nrow(sim_pars)) {

@@

        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "CPU",
            "-lib", "bigsimr"
        )
        dir.create("./results/sims_gamma", showWarnings = FALSE)
        saveRDS(x, file = paste0("./results/sims_gamma/", id, ".rds"))

@@

res <- data.frame()
for (i in 1:nrow(sim_pars)) {
    type <- sim_pars$type[i]
    cores <- sim_pars$cores[i]
    n <- sim_pars$n[i]
    adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
    ## AGS. 17 Oct 2020. cor_bounds failed.
    ## I'll create a minimal working example and open an issue
    ## tmp_bounds <- cor_bounds(margins, type = type)
    ## cor_lo <- tmp_bounds$lower[1,2] + eps
    ## cor_hi <- tmp_bounds$upper[1,2] - eps
    cor_lo <- -0.75
    cor_hi <- 0.95
    cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "CPU",
            "-lib", "bigsimr"
        )
        dir.create("./results/nbinom_sims", showWarnings = FALSE)
        saveRDS(x, file = paste0("./results/nbinom_sims/", id, ".rds"))

>>>>

```{r LoadLib040, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
# reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

res <- data.frame()
dir.create("./results/sims_norm", showWarnings = TRUE, recursive = TRUE)
for (i in 1:nrow(sim_pars)) {
    type <- sim_pars$type[i]
    cores <- sim_pars$cores[i]
    n <- sim_pars$n[i]
    adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
    tmp_bounds <- cor_bounds(margins, type = type)
    cor_lo <- tmp_bounds$lower[1,2] + eps
    cor_hi <- tmp_bounds$upper[1,2] - eps
    cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
    ## rho = cor_seq[1]
    for (rho in cor_seq) {
        Rho <- matrix(rho, 2, 2)
        diag(Rho) <- 1.0
        ## Rho <- convertCor( rho = Rho, from = type, to = 'pearson' )
        time_data <- system.time({
            ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
            ## use rvec to test our software
            x <- rvec(n = n,
                      rho = Rho,
                      margins = margins,
                      type = type,
                      cores = cores)
        })
        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "1CORE",
            "-lib", "bigsimr"
        )
        ## save sims in their own subdirectory
        saveRDS(x, file = paste0("./results/sims_norm/", id, ".rds"))

@@

dir.create("./results/sims_gamma", showWarnings = TRUE, recursive = TRUE)
res <- data.frame()
for (i in 1:nrow(sim_pars)) {

@@

        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "CPU",
            "-lib", "bigsimr"
        )
        saveRDS(x, file = paste0("./results/sims_gamma/", id, ".rds"))

@@

dir.create("./results/nbinom_sims", showWarnings = TRUE, recursive = TRUE)
res <- data.frame()
for (i in 1:nrow(sim_pars)) {
    type <- sim_pars$type[i]
    cores <- sim_pars$cores[i]
    n <- sim_pars$n[i]
    adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
    ## AGS. 17 Oct 2020. cor_bounds failed.
    ## I'll create a minimal working example and open an issue
    ## tmp_bounds <- cor_bounds(margins, type = type)
    ## cor_lo <- tmp_bounds$lower[1,2] + eps
    ## cor_hi <- tmp_bounds$upper[1,2] - eps
    cor_lo <- -0.75
    cor_hi <- 0.95
    cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

        ## Save the sims in case
        id <- paste0(
            "d", 2,
            "-N", n,
            "-c", cores,
            "-r", rho,
            "-Cor", type,
            "-adj", as.character(adjustForDiscrete),
            "-dev", "CPU",
            "-lib", "bigsimr"
        )
        saveRDS(x, file = paste0("./results/nbinom_sims/", id, ".rds"))

### 2020-10-20 050-examples.Rmd
dir create moved outside of loop and set recursive
some directories did not exist, and so building the book would fail

```{r LoadLib050, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

@@

```{r processBRCA, echo=FALSE, eval=FALSE, results='none'}
## full workflow simulating RNA-seq data
allDat <- readRDS( file = "~/Downloads/complete_processed_tcga2stat_RNASeq2_with_clinical.rds" )
lastClinical <- which( names(allDat) == 'tumorsize' )
brca <- allDat[ allDat$disease == "BRCA", (lastClinical+1):ncol(allDat) ]
dim(brca) ## num of genes 20501
## compute the median expression! avoid the 0s
brcaMedian <- apply(brca, 2, median)
## retain top (1-probs)*100% highest expressing genes for illustration
myProb <- 0.99
cutPoint <- quantile( x = brcaMedian, probs = myProb )
## genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
brca <- brca[ , genesToKeep ]
## ncol(brca) / 20501
print(d <- length(genesToKeep))
## save data as rds
saveRDS(brca, './data/brca99.rds')
```

```{r readBRCA, echo=FALSE, eval=TRUE}
## full workflow simulating RNA-seq data
brca <- readRDS(file = './data/brca99.rds') 
```

>>>>

```{r LoadLib050, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
# reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
set.seed(10142020)
```

@@

```{r processBRCA, echo=FALSE, eval=FALSE, results='none'}
## full workflow simulating RNA-seq data
allDat <- readRDS( file = "~/Downloads/complete_processed_tcga2stat_RNASeq2_with_clinical.rds" )
lastClinical <- which( names(allDat) == 'tumorsize' )
brca <- allDat[ allDat$disease == "BRCA", (lastClinical+1):ncol(allDat) ]
dim(brca) ## num of genes 20501
## compute the median expression! avoid the 0s
brcaMedian <- apply(brca, 2, median)
## retain top (1-probs)*100% highest expressing genes for illustration
myProb <- 0.99
cutPoint <- quantile( x = brcaMedian, probs = myProb )
## genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
genesToKeep <- names( brcaMedian ) [ which(brcaMedian >= cutPoint) ]
brca <- brca[ , genesToKeep ]
## ncol(brca) / 20501
print(d <- length(genesToKeep))
## save data as rds
saveRDS(brca, 'data/brca99.rds')
```

```{r readBRCA, echo=FALSE, eval=TRUE}
## full workflow simulating RNA-seq data
brca <- readRDS(file = 'data/brca99.rds') 
```
